---
title: "Clustering load curves"
output:
  rmdformats::html_docco:
    thumbnails: true
    lightbox: true
    gallery: true
pkgdown:
  as_is: true  
---

Along this vignette, the clustering of load curves is presented. Data from a single
building in biggr dataset is used as sample case. Daily load patterns are obtained from
clustering data considering both absolute consumption and temperature

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Load the needed libraries

```{r setup,echo=FALSE}
library(data.table)
library(ggplot2)
library(gridExtra)
library(plotly)
library(padr)
library(htmlwidgets)
library(carrier)
library(biggr)
```

# Load an example dataset

The biggr package has internally preloaded six datasets containing hourly electricity consumption and weather data from six different buildings.

```{r}
data(biggr)
df <- electricity3 %>% calendar_components(localTimeZone = "Europe/Madrid")
df <- cbind(df, do.call(cbind,oce::sunAngle(df$time,latitude=41.5,longitude=1))[,c("azimuth","altitude")])
df <- df %>% filter(!duplicated(df$time))
```

```{r}
# Time series plots
ts_p <- ggplot(
    reshape2::melt( df %>% select(time, Qe, temperature, windSpeed, GHI) %>% pad(), 
                    "time")
  ) + 
  geom_line(
    aes(time,value)
  ) + 
  facet_wrap(~variable, scales = "free_y", ncol=1) +
  theme_bw()
ts_p <- ggplotly(ts_p)
ts_p
```

```{r}
# Scatter plots of the electricity consumption and outdoor temperature
grid.arrange(
  ggplot(aggregate(df[,c("temperature","Qe")],by=list("date"=as.Date(df$localtime)),FUN=mean)) + 
      geom_point(
        aes(temperature, Qe),
        size=0.05
    )
)
```

```{r}
# All daily load curves at once
ggplot(df) + 
  geom_line(
    aes(hour, Qe, group=date),
    alpha=0.05
  ) + xlab("hour of the day")
```

```{r}
# All weekly load curves at once
ggplot(df) + 
  geom_line(
    aes(weekhour, Qe, group=paste(strftime(localtime,"%Y"),strftime(localtime,"%U"))),
    alpha=0.1
  ) + xlab("hour of the week")
```

# Clustering the daily load curves

```{r}
clust <- clustering_dlc(
  data = df,
  consumptionFeature = "Qe", 
  outdoorTemperatureFeature = "temperature", 
  localTimeZone = "Europe/Madrid", 
  kMax = 4, 
  inputVars = c("loadCurves","dailyConsumption"),
  loadCurveTransformation = "absolute",
  nDayParts = 24
)
if("s" %in% colnames(df))
  df <- df %>% select(-s) 
df <- df %>% left_join(clust$dailyClassification)
df <- df[!is.na(df$s),] # Once the classification procedure works, no need for this.
```

```{r}
# All daily load curves depending the patterns detected
p <- ggplotly(
  ggplot(df) + 
    geom_line(
      aes(hour, Qe, group=date, col=s),
      alpha=0.05
    ) + 
    xlab("hour of the day") + 
    facet_wrap(~s) +
    theme_bw()
  )
saveWidget(p, "clustering_dlc.html", selfcontained = T)
p
```

