---
title: "Data transformation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{biggr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(biggr)
library(ggplot2)
library(lubridate)
data(biggr)
```

# Transformation

Multiple helper functions are provided


## Decompose calendar

Decompose the time in date, day of the year, day of the week, day of the
weekend, working day, non-working day, season, month, hour, minute, ...

```{r, echo=FALSE}
create_serie <- function(n, values, timestep = "hours", start = ymd_hms("2020-01-01 00:00:00")) {
  funcs <- list(
    "mins" = minutes,
    "hours" = hours,
    "days" = days,
    "months" = months,
    "years" = years
  )
  func <- funcs[[timestep]]
  return(
    data.frame(
      time = seq(start, start + func(n - 1), by = timestep),
      value = values
    )
  )
}
```

```{r}
# Get time serie calendar properties
start <- ymd_hms("2020-07-01 00:00:00")
testdata <- create_serie(5, rep(1, 5), timestep = "hours", start = start)
calendar_components(testdata, "Europe/Madrid") %>% select(-c(is_holidays))
```

## Calculate deegree raw and degree days

Calculate the difference between outdoor temperature and a base temperature

```{r}
start <- ymd_hms("2020-12-27 00:00:00")
testdata <- create_serie(7, c(26, 27, 28, 29, 30, 31, 32), timestep = "days")
baseTemperature <- 25
newtestdata <- degree_raw(testdata, baseTemperature, "cooling")

print(testdata)
print(newtestdata)
```

```{r}
start <- ymd_hms("2020-12-27 00:00:00")
testdata <- create_serie(7, c(26, 27, 28, 29, 30, 31, 32), timestep = "days")
baseTemperature <- 25
newtestdata <- degree_days(testdata, "Europe/Madrid", baseTemperature, "cooling", "D")

print(testdata)
print(newtestdata)
```

## Normalization

Multiple normalization methods are provided

- Range int
- Daily
- ZScore
- Profile

### Range int normalization

```{r}
testdata <- data.frame(
  value1 = c(0, 0, 25, 25, 100, 100),
  value2 = c(0, 0, 50, 50, 100, 100)
)
newtestdata <- normalize_range(testdata, lower = 0, upper = 1, 0.30, 0.75)

print(testdata)
print(newtestdata)
```

### Daily normalization

```{r, fig.width=7,fig.height=4}
serie <- electricity5 %>%
  mutate(
    time=ymd_hms(time)
  ) %>%
  rename(c("value"="Qe")) %>%
  select(time, value) %>%
  head(24*6)
newserie <- normalize_daily(serie, localTimeZone = "Europe/Madrid")

serie %>% ggplot(aes(x=time, y=value)) + geom_point() + ggtitle("ante")
newserie %>% ggplot(aes(x=time, y=value)) + geom_point() + ggtitle("post")
```

### Zscore normalization

```{r, fig.width=7,fig.height=4}
serie <- electricity5 %>%
  mutate(
    time=ymd_hms(time)
  ) %>%
  rename(c("value"="Qe")) %>%
  select(time, value) %>%
  head(24*6)
newserie <- normalize_zscore(serie)

serie %>% ggplot(aes(x=time, y=value)) + geom_point() + ggtitle("ante")
newserie %>% ggplot(aes(x=time, y=value)) + geom_point() + ggtitle("post")
```

### Profile normalization

```{r, fig.width=7,fig.height=4}
serie <- electricity5 %>%
  mutate(
    time=ymd_hms(time)
  ) %>%
  rename(c("value"="Qe")) %>%
  mutate(temperature=value) %>%
  select(time, value, temperature) %>%
  head(24*31)
newserie <- normalize_load(serie, "Europe/Madrid")

serie %>% ggplot(aes(x=time, y=value)) + geom_point() + ggtitle("ante")
newserie
```
